---
title: "SP_Data_Prep"
author: "Wendel Raymond"
date: "February 12, 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Squid Pop data preperation
This script will pull in raw data and prepare it for analysis and visualization. 

The overall goal is to test if the proporiton of missing/consumed squid baits is affected by sea otter presense/density.

```{r libraries}
library(tidyverse)
library(corrplot)
library(sp)
library(rgdal)
library(rgeos)
library(raster)
library(gdistance)
library(spatstat)
library(sf)
library(bcmaps)

theme_set(theme_classic())
```

```{r functions, echo = FALSE}
st.er <- function(x, na.rm = TRUE) {
  stopifnot(is.numeric(x))
  return(sd(x, na.rm = na.rm)/sqrt(length(x)))
}
```

## Data
We will load the following

1) 2018 Squid pop data
2) 2018 Sea otter density surveys
3) 2018 Shell litter data
4) 2018 Crab data
5) 2018 Fish data

```{r data}
## Squid Pops ##
pops <- read.csv("squid_pops_data_R_4.11.2019MB.csv", header = TRUE, stringsAsFactors = FALSE)
pops <- pops[1:23, ] # remove rows that just have NAs in them
pops$n_fallen_over[is.na(pops$n_fallen_over)] <- 0 # covert n fallen over NA to 0

## Sea otter surveys ##
otts <- readOGR(dsn = "./Spatial_data", layer = "otts_gps_2018_UTM")

## Sites (sea otter survey sites - so really Tiff and Wendel's sites)##
sites <- readOGR(dsn = "./Spatial_data", layer = "seagrass_sites_2018_UTM")

## Shell Litter ##
shell <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Aa4ec0dd9-7c78-480e-8b08-26c64804fe11", method = "libcurl"), header = TRUE, stringsAsFactors = FALSE)

## Crab ##
crb18 <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A3c238a13-4db0-4157-a6e0-62704c09b1ea", method = "libcurl"), stringsAsFactors = FALSE, header = TRUE)
crb18$taxon <- ifelse(crb18$taxon == "Decapod", "Decapoda", crb18$taxon)

crabLW <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A86a12a5b-0528-457b-a313-26f595293b30", method = "libcurl"), stringsAsFactors = FALSE, header = TRUE)

colnames(crb18)[1] <- "site" # fix typo
crb18$string <- as.character(crb18$string) # treat as character not number
crb18$trap <- ifelse(crb18$trap == "crab ", "crab", crb18$trap) # fix typo

## Fish ##
fsh18 <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A0cae20e4-8f8e-427e-b7b9-5d8dd7111ec0", method = "libcurl"), stringsAsFactors = FALSE, header = TRUE)

## POW water ##
h2o.utm <- readOGR(dsn = "E:/wraymond2/My Documents/Graduate School/APECS General/Sea_otter_index_2020/Sea otter survey 2020/POW_shoreline_UTM", layer = "POW_waters_utm_LD")

h2o.latlong <- spTransform(h2o.utm, CRS("+init=epsg:4326"))
```

### Pops
Calculate  proportion consumed for top and bottom baits per site. Presever site metadata.
```{r pops}
## Bottom Baits ##
pops$prop_botts_cons <- pops$full_bottoms / pops$bottom_pop

## Top Baits ##
pops$prop_tops_cons <- pops$full_tops / pops$top_pop

## Quick Plots ##
par(mfrow = c(2, 2))
hist(pops$prop_botts_cons)
hist(pops$prop_tops_cons)
plot(pops$prop_botts_cons, pops$prop_tops_cons)
par(mfrow = c(1, 1))
```

### Sea otter densities
We need to convert the site and otts data into spatial data frames

#### Spatial prep
```{r sea otters}
## Survey Sites ##
sites <- sites %>% 
  filter(grepl('Tiff', Lead))
sites <- sites[order(sites$site_name),]
sites$site <- unique(otts$site)
```

#### Calculate survey area
```{r}
## Calculate density ##
d <- data.frame() # set up blank data frame to append to later
for(i in unique(1:26)){
  s <- sites[i,] # subset sites 
  b <- gBuffer(s, byid = TRUE, width = 3704, quadsegs = 10) # create 3704 meter radius buffer around site
  b <- spTransform(b, CRS("+init=epsg:4326")) # change projection to for next step
  bw <- gIntersection(h2o.latlong, b, byid = TRUE) # clip the buffer area by water so we only have water area left
  bw <- spTransform(bw, CRS("+proj=utm +zone=8 +ellps=GRS80 +towgs84=0,0,0")) # change projection back to UTM
  e <- extent(bw) # save the extent of the polygon for later use
  r <- raster(xmn = e@xmin, xmx = e@xmax, ymn = e@ymin, ymx = e@ymax, ncol = (e@xmax - e@xmin) / 50, nrow = (e@ymax - e@ymin) / 50, crs = "+proj=utm +zone=8 +ellps=GRS80 +towgs84=0,0,0") # create a raster with 50 X 50 meter grid cells. This raster is "empty"
  wr <- rasterize(bw, r, crs = "+proj=utm +zone=8 +ellps=GRS80 +towgs84=0,0,0") # append water polygon data to raster. This raster is "full"
  wr[wr] <- 50 # convert all raster cell values to a value of 50 (because 50 meters)
  wt <- transition(wr, transitionFunction = function(x){50}, directions = 8) # conver raster to transition layer (required for accCost() below)
  wt <- geoCorrection(wt, type = "r", scl = FALSE) # not really sure what this actually does but you have to do it
  sa <- accCost(wt, s) # calculate the cumulative cost distance
  values(sa) <- (values(sa) * 50) # multiply cost values by 50 to reflect cost in terms of actual distance
  sa[values(sa) > 3700] <- NA # Convert cost rasters > 3700 to NA, now the raster will reflect the true survey area
  p <- rasterToPolygons(sa, dissolve = TRUE) # covert raster to polygon
  #p <- spTransform(p, CRS("+proj=utm +zone=8 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs")) # define projection
  a <- (2500 * length(p) / 1000000) # caculate area of the polygon in sq km. 2500 becasue 50*50=2500
  o <- over(p, otts) # identifies surveyed otters that intersect/overlap with the polygon
  o <- as.data.frame(unique(o)) # convert that subset of otters to a data frame
  o$area <- a # append the area of the polygon to that data frame
  o$site.site <- sites@data[i,1] # gives sites that the otter pt is associated with based on site df
  d <- rbind(d, o) # bind this data frame to the master one. -----> now go back to the top and start again!
}
```

#### Clean up calculations
```{r}
## Clean up raw loop output ##
# Convert n_otters to numeric
d$n_otter <- as.numeric(d$n_otter) 

# Remove NA row produced for each site
d <- na.omit(d)

# save area values for latter use larger values if multiple
ar <- unique(d[, c(2, 9)])
ar <- ar[order(ar$site, -abs(ar$area) ), ]
ar <- ar[ !duplicated(ar$site), ]
ar <- rbind(ar, data.frame(site = "Garcia Cove", area = 99999)) # fill in dummy for garcia

# Count total otters are each site survey date
ott.dat <- d %>% 
  group_by(site, replicate) %>% 
  summarise(n = sum(n_otter))

# Create dummy data frame of all sites #
dum <- expand.grid(unique(otts@data$site), unique(ott.dat$replicate))
colnames(dum)[1] <- "site"
colnames(dum)[2] <- "replicate"

# Merge density otters with dum
ott.dat <- merge(ott.dat, dum, by = c("site", "replicate"), all.y = TRUE)
ott.dat[is.na(ott.dat)] <- 0

# append area
ott.dat <- merge(ott.dat, ar, by = "site")

## Actaually calucate density  ##
ott.dat$dens <- ott.dat$n / ott.dat$area

ott.dat.sum <- ott.dat %>%
  group_by(site) %>% 
  summarise(dens_avg = mean(dens))

## EXPORT so you dont have to run the loop everytime ##
write.csv(ott.dat, "sea_otter_density_2018.csv", row.names = FALSE)
```

#### Read in
So you don't have to run the loop every time.
```{r}
ott.dat <- read.csv("sea_otter_density_2018.csv", header = TRUE, stringsAsFactors = FALSE)

ott.dat.sum <- ott.dat %>%
  group_by(site) %>% 
  summarise(dens_avg = mean(dens))
```


### Shell litter
There are a few features form this data that may be of interest. However the most important may be a) proportion of shells with "crab" death estimte and b) proportion of shells with "sea otter" death estimate.

We will count the number of shells in each death estimtate type and then calcualte the proporiton for each site. We will sum across all transects as this will reflect the total predation at the site. Crabs are readonably mobile.
```{r shells}
## Calculate rel. freq (proprtion) ##
shell.site <- shell %>% 
  group_by(site_name, death_estimate) %>% 
  summarise(n_death_est = n()) %>% 
  mutate(freq = n_death_est / sum(n_death_est))

shell.dum <- expand.grid(unique(shell$site_name), unique(shell$death_estimate))
colnames(shell.dum)[1] <- "site_name"
colnames(shell.dum)[2] <- "death_estimate"
shell.dum$n_death_est <- NA
shell.dum$freq <- NA

shell.site <- merge(shell.site, shell.dum, by = c("site_name", "death_estimate"), all.y = TRUE)

shell.site[is.na(shell.site)] <- 0

shell.site <- shell.site[, 1:4]

## Extract only crab crack ##
shell.site.crab <- shell.site %>% 
  filter(death_estimate == "Crab")
```

### Crab
There are a few features form this data that may be of interest. Mean total crab biomass is of primary interest but also the mass by species may be of interest.

#### Calculations
The first thing to do is to add the length - weight conversions to the staring data frame so that we can convert carapace widths the biomass. This will be done by adding the species specific coefficients to the data frame. Note the this conversion is to what is considered sea otter edible biomass. Also note that fish are recorded in this data that might be useful, but from this point on we are going to remove all fishes. 

```{r crabs only}
crb18 <- crb18 %>% 
  filter(taxon == "Decapoda")
```

Now we will add length weight conversion coefficients
```{r adding length to weight conversion}
crb18 <- merge(crb18, crabLW[, c(1, 5:10)], by = "sp_code", all.x = TRUE)
```

Calculate mass
```{r calcualte mass}
crb18$mass_g <- (crb18$a_mean * (crb18$carapace_width_mm)^crb18$b_mean)
range(crb18$mass_g, na.rm = TRUE)
```

```{r}
## Total by site, string ##
crb18_sum <- crb18 %>% 
  group_by(place_name, string) %>% 
  summarize(total_mass = sum(mass_g))

## Add zeros back in ##
sit <- sort(rep(c("2018_L_01", "2018_H_01", "2018_H_02", "2018_H_03", "2018_L_03", "2018_L_02"), 4))
dum <- data.frame(site = sit, string = rep(c("1", "2", "3", "4"), 6))

crb18_sum <- merge(crb18_sum, dum, all.y = TRUE)
crb18_sum[is.na(crb18_sum)] <- 0

## Summary ##
crb18_site <- crb18_sum %>% 
  group_by(place_name) %>% 
  summarise(crb_mass_avg = mean(total_mass),
            srb_mass_se = st.er(total_mass))
```

### Fish
There are a few features form this data that may be of interest. Total fish biomass is of primary interest but also the mass by species may be of interest.

```{r}
## Total Site Fish Biomass ##
fsh18_site <- fsh18 %>% 
  filter(place_name != "North Fish Egg Island") %>% 
  filter(bay_sample == "A") %>% 
  group_by(place_name) %>% 
  summarise(fsh_mass = sum(mass_g))

## Biomass by sp by Site ##
fsh18_sp_site <- fsh18 %>%
  filter(place_name != "North Fish Egg Island") %>% 
  filter(bay_sample == "A") %>% 
  group_by(place_name, sp_code) %>% 
  summarise(fsh_mass = sum(mass_g),
            species_common = unique(species_common),
            species_scientific = unique(species_scientific))
```

## Curate data
The goal is to edit the dataframe about so that there is consistent site names and 1:1 data coverage for as many sites as possible. 

There are essentially two types of data. (1) Data that is replicated across all sites and (2) data that only exists at the 'sea otter region' scale. So...

1) Sea otter density, pop, and shell data can all be combined into a single dataframe
2) Crab and fish data can be combined into a single data frame.

Note that summary information from dataframe 1, like average sea otter density, could be added to dataframe 2, and vise versa. 

### Dataframe 1 - Pops, Otts, n Shells

Manually assign pops to sea otter/shell sites. This will be done by creating a new column in the pops dataframe with 
new site names. I did this by looking at ArcGIS and deciding what survey site 'should' be assigned to what pop site.

#### Sea otter density
```{r}
## Make list of new names that will be easy to combine ##
pop.new.names <- c("Blanquizal Bay", "Garcia Cove", "S Wadleigh Island", "S Fish Egg Island", "Shinaku Inlet",
                   "N Fish Egg Island", "Sand Lance Cove/ San Fernando", "Big Tree Bay", "Sukkwan Narrows", "Inner Hydaburg", "Goat Mouth Inlet", "Natzuhini Bay 2", "North Pass 2", "Inner Hydaburg", "Soda Bay", "Farallon Bay", "Mushroom Island", "Big Clam Bay", "Goat's Nose", "Inner Hydaburg", "Nossuk Bay Complex", "Nossuk Bay Complex", "Nossuk Bay Complex")

## Make new column in pops df ##
pops$site_comb <- pop.new.names

## Calcualte density for synthetic sites ##
# Inner Hydaburg combines Sukkwan, Mushroom, and the two Natzuhini sites #
dens_InnerHyd <- sum(ott.dat.sum$dens_avg[ott.dat.sum$site == "Sukkwan Narrows"], ott.dat.sum$dens_avg[ott.dat.sum$site == "Mushroom Island"], ott.dat.sum$dens_avg[ott.dat.sum$site == "Natzuhini Bay 2"], ott.dat.sum$dens_avg[ott.dat.sum$site == "Natzuhini Bay 3"]) / 4

# Nossuk Bay Complex combines the two Nossuk bay sites #
dens_NossBC <- sum(ott.dat.sum$dens_avg[ott.dat.sum$site == "Nossuk Bay 2"], ott.dat.sum$dens_avg[ott.dat.sum$site == "Nossuk Bay 3"]) / 2

## Bind these new synth sites ##
otter.new.dens <- rbind(data.frame(site = c("Inner Hydaburg", "Nossuk Bay Complex"), 
                                   dens_avg = c(dens_InnerHyd, dens_NossBC)))


ott.dat.sum.comb <- rbind(ott.dat.sum, otter.new.dens)

## Merge with pop data ##
pop.redu <- pops %>% 
  filter(site_comb != "Sand Lance Cove/ San Fernando" & site_comb != "Farallon Bay" & site_comb != "Goat's Nose")

## Make Dataframe 1 ##
df1 <- merge(pop.redu, ott.dat.sum.comb, by.x = "site_comb", by.y = "site", all.x = TRUE)
df1[11, 2] <- "Natzuhuni"
```

#### Shells
```{r}
shell_NossBC <- sum(shell.site.crab$freq[shell.site.crab$site_name == "Nossuk Bay b"], shell.site.crab$freq[shell.site.crab$site_name == "Nossuk Bay c"]) / 2

shell_InnerHyd <- sum(shell.site.crab$freq[shell.site.crab$site_name == "Mushroom Island a"], shell.site.crab$freq[shell.site.crab$site_name == "Natzuhini Bay b"], shell.site.crab$freq[shell.site.crab$site_name == "Natzuhini Bay c"], shell.site.crab$freq[shell.site.crab$site_name == "Sukkwan Narrows a"]) / 4

pop.sites <- df1$site

shell.fq <- c(0.1026, 0.0388, 0.0256, 0.0106, 0.0882, shell_InnerHyd, shell_InnerHyd, shell_InnerHyd, 0.0526, 0.0000, 0.0690, 0.1154, 0.0152, shell_NossBC, shell_NossBC, 0.0168, 0.1471, 0.0000, 0.0982, 0.0119)

## Bind these new synth sites ##
shell.site.crab.new <- data.frame(site = pop.sites, prop_crbshl = shell.fq)

## Merge with df1 ##
df1 <- merge(df1, shell.site.crab.new)
```

### Dataframe 2 - crabs and Fish

Add a high/low sea otter indicators in crab and fish dataframes. Then add some summarized data from df1

#### Crabs
```{r}
crb18_site$sea_otter_level <- c("high", "high", "low", "low", "high", "low")
```

#### Fish
```{r}
fsh18_site[1,1] <- "Gaktu Cove"
fsh18_site[3,1] <- "Natzuhini"

df2 <- merge(crb18_site, fsh18_site)
```

#### Add summarized df1 data
```{r}
df1.sum <- df1 %>% 
  group_by(sea_otter_level) %>% 
  summarize(prop_botts_cons_avg = mean(prop_botts_cons),
            prop_botts_cons_se = st.er(prop_botts_cons),
            ott_dens_avg = mean(dens_avg),
            ott_dens_se = st.er(dens_avg),
            prop_crbshl_avg = mean(prop_crbshl),
            prop_crbshl_se = st.er(prop_crbshl))

df2 <- merge(df2, df1.sum)
```

## Export Final Data
```{r}
write.csv(df1, "pops_master_allsites.csv", row.names = FALSE)
write.csv(df2, "pops_master_region.csv", row.names = FALSE)
```

